---
title: "Predictive Maintenance using the AI4I 2020 Dataset"
urlcolor: Blue
author: Trevor King
date: 2024-05-14
output:
  pdf_document:
    dev: png
    toc: yes
    toc_depth: 3
  html_notebook:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: true
    toc_depth: '3'
    df_print: paged
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)
```
```{r getlibs,echo=FALSE,message=FALSE,warning=FALSE}
library(readr)
library(knitr)
library(tidyverse)
library(caret)
library(GGally)
```

# Introduction
The dataset was downloaded from Kaggle and is modeled after an existing milling machine.
The dataset has 10000 rows with 14 columns. The 14 columns contain information about the product ID and type as well as temperature, tourque, rpm, tool wear, and 6 columns for failure data.


ai4i2020.csv


The Fields are as follows:

1.    UID
2.    Product ID Unique values (Prefixed with L - Low, M - Medium, and H - High)
3.    Product Type product type L, M or H from 2 above
4.    air temperature [K]
5.    Process temperature [K]
6.    Rotational Speed [RPM]
7.    Torque [Nm]
8.    Tool wear [min]
9.    Target
10.    Machine failure

# Methods/Analysis

## Exploratory Data Analysis

A preview of the table structure.


```{r intro_code}
# Read in the CSV file
FN <- "C:/Users/kingt/Documents/CapstonePredictiveMaintenance/data/predictive_maintenance.csv"

x<-spec_csv(FN,col_types = cols())

read_csv(FN)%>%apply(., 2, function(x) {length(unique(x))})%>%kable(col.names = c('Length'))
```

Note that the Type only has 3 distinct values which makes it a good candidate for factorization.

And the last five columns only have two values, 0 and 1, we can bring these in as logical

```{r EDA_code}

# Import the data with defined field types
maintlog<-read_csv(FN,col_types = 'icfddiddlf')

# Display the 1st 10 rows
maintlog%>%head()%>%kable()
```

```{r fg1}
# Remove the squre brackets from the vars
names(maintlog)<-names(maintlog)%>%sub("\\s\\[.{1}.*\\]$", "", .)

# Replace the spaces with '_' in the column names (Easier to work with)
names(maintlog)<-names(maintlog)%>%gsub("\\s", "_", .)

value_features<-names(maintlog)[4:8]

# Create a pair plot of 
vf<-maintlog[,4:10]
# vf%>%pairs(col = 'blue', #modify color
#      main = 'Features vs. Machine Failure')
```

### Numeric Variables

Quantitative variables

Air Temperature

```{r fg2}
vf%>%qplot(Air_temperature, geom ="histogram", data = .,
colour = I("#2EA7CE"), fill = I("#8E87CE"),
main = "Air Temperature", bins=50)

summary(vf$Air_temperature)

```

Process Temperature
```{r fg3}
vf%>%qplot(Process_temperature, geom ="histogram", data = .,
colour = I("#2EA7CE"), fill = I("#8E87CE"),
main = "Process Temperature", bins=50)

vf$Process_temperature%>%summary()

```

Rotational Speed
```{r fg4}
vf%>%qplot(Rotational_speed, geom ="histogram", data = .,
colour = I("#2EA7CE"), fill = I("#8E87CE"),
main = "Rotational Speed", bins=50)

vf$Rotational_speed%>%summary()

```

Torque
```{r fg5}
vf%>%qplot(Torque, geom ="histogram", data = .,
colour = I("#2EA7CE"), fill = I("#8E87CE"),
main = "Torque", bins=50)

vf$Torque%>%summary()
```

Tool wear
```{r fg6}
vf%>%qplot(Tool_wear, geom ="histogram", data = .,
colour = I("#2EA7CE"), fill = I("#8E87CE"),
main = "Tool wear", bins=50)

vf$Tool_wear%>%summary()

```

```{r fig7}

maintlog$Type <- factor(maintlog$Type, levels=c('L', 'M', 'H'))

maintlog$Type%>%table()%>%kable()

maintlog%>%ggplot(aes(Type, fill = Failure_Type)) + 
  geom_histogram(stat="count") + ggtitle("Distribution of Machine Types")


```

```{r fg8}
# Air Temp vs. Proccess Temp
p1 <- ggplot(vf)+geom_point(aes(x = Air_temperature, y = Process_temperature,colour = Failure_Type), na.rm = TRUE)

# Air Temp vs. Rotation Speed
p2 <- ggplot(vf)+geom_point(aes(x = Air_temperature, y = Rotational_speed,colour = Failure_Type), na.rm = TRUE)

# Air Temp vs. Torque
p3 <- ggplot(vf)+geom_point(aes(x = Air_temperature, y = Torque,colour = Failure_Type), na.rm = TRUE)

# Air Temp vs. Tool wear
p4 <- ggplot(vf)+geom_point(aes(x = Air_temperature, y = Tool_wear,colour = Failure_Type), na.rm = TRUE)

# Process temp vs. Rotation Speed
p5 <- ggplot(vf)+geom_point(aes(x = Process_temperature, y = Rotational_speed,colour = Failure_Type), na.rm = TRUE)

# Process temp vs. Torque
p6 <- ggplot(vf)+geom_point(aes(x = Process_temperature, y = Torque,colour = Failure_Type), na.rm = TRUE)

# Process temp vs. Tool wear
p7 <- ggplot(vf)+geom_point(aes(x = Process_temperature, y = Tool_wear,colour = Failure_Type), na.rm = TRUE)

# Rotation Speed vs. Torque
p8 <- ggplot(vf)+geom_point(aes(x = Rotational_speed, y = Torque,colour = Failure_Type), na.rm = TRUE)

# Rotation Speed vs. Tool wear
p9 <- ggplot(vf)+geom_point(aes(x = Rotational_speed, y = Tool_wear,colour = Failure_Type), na.rm = TRUE)

# Torque vs. Tool wear
p10 <- ggplot(vf)+geom_point(aes(x = Torque, y = Tool_wear,colour = Failure_Type), na.rm = TRUE)


# observe how these features interact under different failure conditions. 
gridExtra::grid.arrange(p1,p2,p3,p4,p5)
gridExtra::grid.arrange(p6,p7,p8,p9,p10)


```
## Creating the training, testing and validation sets

```{r}
# Need to check the R version since the syntax is different for R version>3.6
# Set the random seed value
if(base::getRversion()>'3.6'){ set.seed(1, sample.kind="Rounding") }else{set.seed(1) }

# Partition the data by creating an index where 10% is for the hold out data and
# 90% for the remain
test_index <- createDataPartition(y = maintlog$UDI, times = 1, p = 0.1, list = FALSE)
train_set <- maintlog[-test_index,]
test_set <- maintlog[test_index,]

plt<-table(train_set$Type)/length(train_set$Type)*100
barplot(plt, col = c(I("#a7ce2e"),I("#8EA7CE"),I("#CE0000")),main = "Percentages by Type")

plt<-table(train_set$Target)/length(train_set$Target)*100
barplot(plt, col = c(I("#a7ce2e"),I("#CE0000")),main = "Percentages by Type")

train_qvar <- train_set[, - 2]
Target<- as.factor(train_set$Target)
ggpairs(train_qvar, aes(colour=Target))


```

```{r}

train_set[,-1]%>% 
  ggplot(aes(Rotational_speed, Torque, fill = Target, color=Failure_Type)) + 
  geom_point(show.legend = FALSE) + 
  stat_ellipse(type="norm", lwd = 1.5)
# Logistic regression

# train_qda <- train(Failure_Type ~ ., method = "qda", data = train_set[,-1])


```
 
# Results

# Conclusion

# references 
S. Matzka, "Explainable Artificial Intelligence for Predictive Maintenance Applications," 2020 Third International Conference on Artificial Intelligence for Industries (AI4I), 2020, pp. 69-74, doi: 10.1109/AI4I49448.2020.00023.
